Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-05-04 doc/thesis/results.tex
Moar

\subsubsection*{Classifiers for parameters based on tree shape}

@@

The accuracy of grid search estimates largely paralleled that of the \gls{kSVR}
classifiers. \Cref{fig:gridest} shows point estimates and 95\% highest density
intervals for each of the \gls{BA} parameters, for one replicate experiment
with 500-tip trees. Plots showing the point estimates for all replicates can be
found in \cref{fig:gridptalpha,fig:gridptI,fig:gridptm,fig:gridptN}. For all
parameters except $m$, the error of point estimates was negatively correlated
with the number of sampled tips in the tree (for
\gls{alpha}, \gls{I}, and \gls{N} respectively: Spearman's $\rho$ = 
    \ensuremath{-0.22},
    \ensuremath{-0.51},
    \ensuremath{-0.16};
$p$-values
    $4\!\times\!10^{-4}$,
    ${<}10^{-5}$,
    $0.01$).
The highest density intervals obtained for all parameters were extremely wide,
occcupying $>$90\% of the grid in all cases (\cref{fig:gridest}).

The \gls{alpha} parameter was the most accurately estimated, with point
estimates having an average deviation of 
    0.14
from the true value, on a grid from 0 to 2. The error was negatively correlated
with the true value of \gls{alpha} 
    (Spearman's $\rho$ = \ensuremath{-0.26},
     $p = 10^{-5}$),
although the relationship was clearly nonlinear (\cref{fig:gridest,fig:gridalpha}).
The accuracy was highest for the test value \gls{alpha} = 1.25
    (mean error 0.02)
which exhibited markedly different behaviour than the other values in terms of
the distribution of kernel scores along the grid (\cref{fig:gridalpha}). In
particular, there was a very pronounced peak in scores around the true value,
in contrast to most other values where the scores were flat around the true
value. The peak was also observed to a lesser degree for \gls{alpha} = 1. The
average absolute error of the point estimates for \gls{I} was 
    310,
on a grid of 500 to 5000, and the errors were not significantly correlated with
the true value of \gls{I}. Kernel score distributions for all test values
exhibited a similar rounded shape (\cref{fig:gridI}).

The average error for \gls{m} was
    1.31,
on a grid from 1 to 6; this error was positively correlated with increasing
\gls{m}
    (Spearman's $\rho$ = 0.25,
     $p = 6\!\times\!10^{-4}$).
This positive correlation was apparently due to the much lower error for
\gls{m} = 1 and 2 than for the other $m$ values 
    (mean errors 
    0.93 
    for $m \leq 2$ vs. 
    1.49 
    for $m > 2$, \cref{fig:gridm}).
The value $m = 1$ causes the network to take on a distinct shape relative to
higher \gls{m} values, namely a tree (\textit{i.e.} there are no cycles,
see \cref{subsec:treeshape}). The average error for \gls{N} was 
    2419,
on a grid from 1000 to 15000, and was positively correlated with the true value
of \gls{N}
    (Spearman's $\rho$ = 0.43,
    $p\ {<}10^{-5}$). 
Most of the kernel score distributions were extremely flat, except for the
value $N = 1000$ which exhibited a pronounced peak around the true value
(\cref{fig:gridN}).

@@

We used \software{netabc} to estimate the parameters of the \gls{BA} model on
simulated trees where the true parameter values were known. Point estimates for
each parameter are shown in \cref{fig:abcptm2} for the simulations with \gls{m} =
2. The results for the other values of \gls{m} were similar
(\cref{fig:abcptm3,fig:abcptm4}). The median [IQR] absolute error of estimates of
\gls{alpha} across all simulations was
    0.08 
    [0.05-0.17].
The accuracy of the estimates was not significantly different between values of
$m$ or $I$ (both one-way ANOVA,
    $p$ = 0.05
and 
    0.07),
although the errors when the true value of \gls{alpha} was zero were
significantly greater than the other values 
    (Wilcoxon rank-sum test, $p = 8\!\times\!10^{-3}$).
The error in the estimated value of $I$ was
    395 
    [207-683].
Errors were significantly higher for $\alpha \geq 1$
    (Wilcoxon rank-sum test, $p = 8\!\times\!10^{-3}$)
and for $I$ = 2000
    ($p = {<}10^{-5}$),
but not for any values of $m$ (one-way ANOVA). The $m$ parameter was estimated
correctly in only
    27 \%
of simulations, little better than random guessing. The true values of $m$ and
$I$ did not significantly affect the error (one-way ANOVA). Finally, the total
number of nodes \gls{N} was consistently over-estimated by about a factor of
two
    (error 5987 
    [2060 - 
     7999]).
No other parameters influenced the accuracy of the $N$ estimates (one-way
ANOVA).

\begin{figure}
  \includegraphics{abc-point-estimate-m2}
  \caption[\Acrlong{MAP} point estimates for \gls{BA} model parameters obtained
    by running \software{netabc} on simulated data, for simulations with $m = 2$.] 
  {
    \Acrlong{MAP} point estimates for \gls{BA} model parameters obtained by         
    running \software{netabc} on simulated data. Values shown are for               
    simulations with \gls{m} = 2. Dashed lines indicate true values. (A)            
    Estimates of \gls{alpha} and \gls{I} which were varied in these simulations  
    against known values. (B) Estimates of \gls{m} and \gls{N} which were held   
    fixed in these simulations at the values \gls{m} = 2 and \gls{N} = 5000. 
  }
  \label{fig:abcptm2}
\end{figure}

\Cref{fig:abcex} shows the \gls{ABC} approximation to the posterior
distribution on the \gls{BA} parameters for one simulation. Equivalent plots
for one replicate simulation with each combination of parameters can be found
in \cref{fig:0.0-1000-2-5000-0,fig:0.5-1000-2-5000-0,fig:1.0-1000-2-5000-0,fig:1.5-1000-2-5000-0,fig:0.0-2000-2-5000-0,fig:0.5-2000-2-5000-0,fig:1.0-2000-2-5000-0,fig:1.5-2000-2-5000-0,fig:0.0-1000-3-5000-0,fig:0.5-1000-3-5000-0,fig:1.0-1000-3-5000-0,fig:1.5-1000-3-5000-0,fig:0.0-2000-3-5000-0,fig:0.5-2000-3-5000-0,fig:1.0-2000-3-5000-0,fig:1.5-2000-3-5000-0,fig:0.0-1000-4-5000-0,fig:0.5-1000-4-5000-0,fig:1.0-1000-4-5000-0,fig:1.5-1000-4-5000-0,fig:0.0-2000-4-5000-0,fig:0.5-2000-4-5000-0,fig:1.0-2000-4-5000-0,fig:1.5-2000-4-5000-0}. \Gls{HPD} intervals around \gls{alpha} and \gls{I} were
narrow relative to the region of nonzero prior density, whereas the intervals
for $m$ and \gls{N} were widely dispersed. \Cref{tab:abchpd} shows point
estimates and 95\% \gls{HPD} intervals averaged over all simulations.

\begin{figure}
  \includegraphics{{abc-posterior/1.0_1000_2_5000_0}.pdf}
  \caption[Approximate marginal posterior distributions of BA model parameters
      obtained by applying \textit{netabc} to a simulated transmission tree
      with values $\alpha$ = 1.0, $I$ = 1000, $m$ = 2, and $N$ = 5000.]
    {
        Approximate marginal posterior distributions of BA model parameters
        obtained by applying \textit{netabc} to a simulated transmission tree
        with BA parameter values $\alpha$ = 1.0, $I$ = 1000, $m$ = 2, and $N$ =
        5000. Vertical dashed lines indicate true values. Shaded areas are 95\%
        highest posterior density intervals. $x$-axes indicate regions of
        nonzero prior density.
    }
  \label{fig:abcex}
\end{figure}

\begin{table}
    \centering
    \input{\tablepath/abc-hpd}
    \caption[
        Maximum \textit{a priori} estimates and 95\% highest posterior density
        (HPD) interval boundaries for \gls{BA} model parameters estimated with
        \software{netabc}, averaged over simulated transmission trees.
    ]
    {
        Maximum \textit{a priori} estimates and 95\% highest posterior density
        (HPD) interval boundaries for \gls{BA} model parameters estimated with
        \software{netabc}, averaged over simulated transmission trees.
    }
    \label{tab:abchpd}
\end{table}

To test the effect of model misspecification, we simulated one network where
the nodes exhibited heterogeneous preferential attachment power (half 0.5, the
other half 1.5), with $m$ = 2, $N$ = 5000, and $I$ = 1000. The MAP [95\%
HPD] estimates for each parameter were: 
$\alpha$, 
  1.1 
  [0.6 -
   1.16];
$I$,
  1091 
  [662 -
   4455];
$m$,
  3 
  [1 -
   5];
$N$,
  12566 
  [3733 -
   14945].
The approximate posterior distributions for this simulation are shown in
\cref{fig:mixed}. To test the effect of sampling bias, we sampled one
transmission tree in a peer-driven fashion, where the probability to sample a
node was twice as high if one of its peers had already been sampled. The
parameters for this experiment were $N$ = 5000, $m$ = 2, $\alpha$ = 0.5, and
$I$ = 2000. The estimated values were
$\alpha$, 
  0.37 
  [0 -
   0.62];
$I$,
  2407 
  [1423 -
   3811];
$m$,
  3 
  [2 -
   5];
$N$,
  13811 
  [2624 -
   14647].
The approximate posterior distributions are shown in \cref{fig:peerdriven}. Both
of these results were in line with estimates obtained on other simulated
datasets (\cref{tab:abchpd}), although the estimate of peer-driven sampling for
$\alpha$ was somewhat lower than typical.

@@

>>>>

\subsubsection*{Classifiers for BA model parameters based on tree shape}

@@

The accuracy of grid search estimates largely paralleled that of the \gls{kSVR}
classifiers. \Cref{fig:gridest} shows point estimates and 95\% highest density
intervals for each of the \gls{BA} parameters, for one replicate experiment
with 500-tip trees. Plots showing the point estimates for all replicates can be
found in \cref{fig:gridptalpha,fig:gridptI,fig:gridptm,fig:gridptN}. For all
parameters except $m$, the error of point estimates was negatively correlated
with the number of sampled tips in the tree (for
\gls{alpha}, \gls{I}, and \gls{N} respectively: Spearman's $\rho$ = 
    \ensuremath{-0.22},
    \ensuremath{-0.51},
    \ensuremath{-0.16};
$p$-values
    $4\!\times\!10^{-4}$,
    ${<}10^{-5}$,
    $0.01$).
The highest density intervals obtained for all parameters were extremely wide,
occcupying $>$75\% of the grid in all cases (\cref{fig:gridest}).

The \gls{alpha} parameter was the most accurately estimated, with point
estimates having an average deviation of 
    0.14
from the true value, on a grid from 0 to 2. The error of point estimates varied
significantly between true values of \gls{alpha}
    (one-way \gls{ANOVA}, $p {<}10^{-5}$). In
particular, errors were lower for the values \gls{alpha} = 1.0 and 1.25 than
for the other values
    (average errors 
    0.03
    for \gls{alpha} = 1.0 or 1.5 vs.
    0.17
    for \gls{alpha} $\neq$ 1.0 or 1.5),
and this difference was significant
    (Wilcoxon rank-sum test, $p {<}10^{-5}$,
     \cref{fig:gridptalpha}).
These two values exhibited different qualitative behaviour than the other
values in terms of the distribution of kernel scores along the grid
(\cref{fig:gridalpha}). In particular, there was a pronounced peak in scores
around the true value, in contrast to the other values where the scores were
flat around the true value. The effect was most obvious for the value
\gls{alpha} = 1.25.

The average absolute error of the point estimates for \gls{I} was 
    310 individuals,
on a grid of 500 to 5000, and these errors differed between true values of
\gls{I}
    (one-way \gls{ANOVA}, $p =0.001$).
The errors for $2000 \leq I \leq 3000$ were higher than those for the other
values
    (average errors
     430
     for $2000 \leq I \leq 3000$ vs.
     250
     for $I < 2000$ or $I > 3000$),
and this difference was significant
    (Wilcoxon rank-sum test, $p =6\!\times\!10^{-4}$,
     \cref{fig:gridptI}).
Kernel score distributions for all test values exhibited a similar rounded
shape (\cref{fig:gridI}).

The average error for \gls{m} was
    1.31 edges per vertex,
on a grid from 1 to 6. The error varied significantly between the true values
of \gls{m} 
    (one-way \gls{ANOVA}, $p {<}10^{-5}$).
Errors for the value \gls{m} = 1 were lower than the other values
    (average errors
    0.1
    for $m = 1$ vs.
    1.55
    for $m > 1$),
and this difference was significant
    (Wilcoxon rank-sum test, $p {<}10^{-5}$,
     \cref{fig:gridptm}).
The value $m = 1$ causes the network to take on a distinct shape relative to
higher \gls{m} values, namely a tree (\ie there are no cycles, see
\cref{subsec:treeshape}). The kernel score distribution had a peak at $m = 1$
when this was the true value, and a valley at $m = 1$ when the true value of
$m$ was greater that 1 (\cref{fig:gridm}).

The average error for \gls{N} was 
    2419 individuals,
on a grid from 1000 to 15000, and was varied significantly with the true value
of \gls{N}
    (one-way \gls{ANOVA}, $p {<}10^{-5}$).
The errors were lower for $N \leq 3000$
    (average errors
     740
     for $N \leq 3000$ vs.
     2979
     for $N > 3000$),
and this diffrence was significant
    (Wilcoxon rank-sum test, $p {<}10^{-5}$,
     \cref{fig:gridptN}).
The kernel score distribution had a peak at $N = 1000$ when this was the true
value, and a valley there otherwise (\cref{fig:gridN}). Except for this valley,
the distributions were flat for $N > 3000$.

@@

\Cref{fig:abcpt} shows \gls{MAP} point estimates of the BA model parameters
obtained with kernel-ABC on simulated data. The estimates shown correspond only
to the simulations where the $m$ parameter was set to 2, however the results
for $m = 3$ and $m = 4$ were similar (\cref{fig:abcptm3,fig:abcptm4}). Average
boundaries of 95\% HPD intervals are given in \cref{tab:abchpd}.

The accuracy of the parameter estimates obtained with kernel-ABC
paralleled the results from the \gls{kSVR} classifier. Of the four parameters,
$\alpha$ was the most accurately estimated, with point estimates having a
median [IQR] absolute error of 
    0.08 
    [0.05 - 
    0.17].
The errors when the true value of $\alpha$ was zero were significantly greater
than those for the other values 
    (Wilcoxon rank-sum test, $p$ = $0.0078$).
Errors in estimating $\alpha$ also varied with the true value of $m$ just at
the threshold of statistical significance
    ($p 0.05$),
but did not vary across the true values of $N$ or $I$ (both one-way ANOVA).
Estimates for $I$ were relatively accurate, with point estimate errors of
    395 
    [207 - 
    683] individuals.
These errors were significantly higher when the true value of $\alpha$ was
at least 1
    (Wilcoxon rank-sum test, $p$ = $0.0077$)
and when the true value of $I$ was 2000 ($p < 10^{-5}$). The true value of $m$
did not affect the estimates of $I$ (one-way ANOVA).

The $m$ parameter was estimated correctly in only
    27 \%
of simulations, barely better than random guessing. The true values of the
other parameters did not significantly affect the estimates of $m$ (both
one-way ANOVA). Finally, the total number of nodes $N$ was consistently
over-estimated by about a factor of two
    (error 5987 
    [2060 - 
     7999] individuals).
No parameters influenced the accuracy of the $N$ estimates (all one-way ANOVA).

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{abc-point-estimate-m2}
  \vspace{6pt}
  \caption{
    Point estimates of BA model parameters obtained by running kernel-ABC
    on simulated phylogenies without training, for simulations with $m = 2$.
    Dotted lines indicate true values, and limits of the $y$-axes are regions
    of uniform prior density. (A) Estimates for $\alpha$ and $I$ against their
    true values in simulations. (B) Estimates for $m$ and $N$, which were held
    fixed in these simulations, against true values of $\alpha$.
  }
  \label{fig:abcpt}
\end{figure}

\begin{table*}[ht]
  \centering
  \input{\tablepath/abc-hpd.tex}
  \caption{
      Average maximum \textit{a posteriori} point estimates and 95\% highest
      posterior density (HPD) interval widths for BA model parameter estimates
      obtained with kernel-ABC. Three transmission trees were simulated under
      each combination of the listed parameter values, and the parameters were
      estimated with kernel-ABC without training.
  }
  \label{tab:abchpd}
\end{table*}

The dispersion of the ABC approximation to the posterior also varied between
the parameters, with narrower HPD intervals for the parameters with more
accurate point estimates (\cref{tab:abchpd}). \Cref{fig:abcex} shows
the distributions for for one simulation. Equivalent plots for one replicate
simulation with each studied parameter combination can be found in
\cref{fig:0.0-1000-2-5000-0,fig:0.5-1000-2-5000-0,fig:1.0-1000-2-5000-0,fig:1.5-1000-2-5000-0,fig:0.0-2000-2-5000-0,fig:0.5-2000-2-5000-0,fig:1.0-2000-2-5000-0,fig:1.5-2000-2-5000-0,fig:0.0-1000-3-5000-0,fig:0.5-1000-3-5000-0,fig:1.0-1000-3-5000-0,fig:1.5-1000-3-5000-0,fig:0.0-2000-3-5000-0,fig:0.5-2000-3-5000-0,fig:1.0-2000-3-5000-0,fig:1.5-2000-3-5000-0,fig:0.0-1000-4-5000-0,fig:0.5-1000-4-5000-0,fig:1.0-1000-4-5000-0,fig:1.5-1000-4-5000-0,fig:0.0-2000-4-5000-0,fig:0.5-2000-4-5000-0,fig:1.0-2000-4-5000-0,fig:1.5-2000-4-5000-0}. HPD intervals around $\alpha$ and $I$ were often narrow
relative to the region of nonzero prior density, whereas the intervals for $m$
and $N$ were more widely dispersed.

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{abc-posterior-example}
  \vspace{6pt}
  \caption{
    Marginal posterior distributions of BA model parameters estimated
    with kernel-ABC for a single simulated transmission tree. Dotted
    lines and shaded polygon indicate true values.
  }
  \label{fig:abcex}
\end{figure}

To test the effect of model misspecification, we simulated one network where
the nodes exhibited heterogeneous preferential attachment power (half 0.5, the
other half 1.5), with $m$ = 2, $N$ = 5000, and $I$ = 1000. The MAP [95\%
HPD] estimates for each parameter were: 
$\alpha$, 
  1.09 
  [0.6 -
   1.16];
$I$,
  1097 
  [662 -
   4455];
$m$,
  3 
  [1 -
   5];
$N$,
  12766 
  [3367-
   14945].
The approximate posterior distributions for this simulation are shown in
\cref{fig:mixed}. To test the effect of sampling bias, we sampled one
transmission tree in a peer-driven fashion, where the probability to sample a
node was twice as high if one of its peers had already been sampled. The
parameters for this experiment were $N$ = 5000, $m$ = 2, $\alpha$ = 0.5, and
$I$ = 2000. The estimated values were
$\alpha$, 
  0.1 
  [0.02 -
   0.63];
$I$,
  2399 
  [1419 -
   3742];
$m$,
  3 
  [2 -
   5];
$N$,
  9338 
  [2881 -
   14647].
The approximate posterior distributions are shown in \cref{fig:peerdriven}. Both
of these results were in line with estimates obtained on other simulated
datasets (\cref{tab:abchpd}), although the estimate of peer-driven sampling for
$\alpha$ was somewhat lower than typical.

%<<abc_glm, include=FALSE>>=
%    source("global.R")
%    options(scipen=-1, digits=2)
%@
%
%<<alpha_glm, include=FALSE>>=
%    # alpha_error is influenced by alpha and m but not I
%    stopifnot(alpha.glm[Parameter == "alpha", min(p) < 0.05])
%    stopifnot(alpha.glm[Parameter == "m", min(p) < 0.05])
%    stopifnot(alpha.glm[Parameter == "I", min(p) > 0.05])
%
%    # alpha_error is correlated with true_alpha
%    alpha.test <- d[,cor.test(alpha_error, true_alpha, method="spearman")]
%    stopifnot(alpha.test$p.value < 0.05)
%
%    # alpha_error is not correlated with true_m
%    alpha.m.cor <- d[,cor.test(alpha_error, true_m)]
%    stopifnot(alpha.m.cor$p.value < 0.05)
%@
%
%We used \software{netabc} to estimate the parameters of the \gls{BA} model on
%simulated trees where the true parameter values were known. Point estimates for
%each parameter are shown in \cref{fig:abcptm2} for the simulations with \gls{m} =
%2. The results for the other values of \gls{m} were similar
%(\cref{fig:abcptm3,fig:abcptm4}). The median [IQR] absolute error of estimates
%of \gls{alpha} across all simulations was
%    d[,median(alpha_error)] 
%    [d[,quantile(alpha_error, 0.25)]-d[,quantile(alpha_error, 0.75)]].
%\Gls{GLM} analysis indicated that the true values of both \gls{alpha} and
%\gls{m} had significant effects on the error in estimated \gls{alpha}
%    ($p$ values $pp(alpha.glm[Parameter == "alpha", min(p)], eq=FALSE)$
%     $pp(alpha.glm[Parameter == "m", min(p)], eq=FALSE)$,
%     \cref{tab:glmalpha}),
%but the true value of \gls{I} did not. There was a significant negative
%correlation between the true value of \gls{alpha} and the error
%    (Spearman's $\rho$ = round(alpha.test$estimate, 2),
%     $p pp(alpha.test$p.value, eq=TRUE)$).
%There was a negative correlation between \gls{m} and the error in \gls{alpha}
%    (Spearman's $\rho$ = round(alpha.m.cor$estimate, 2),
%    $p pp(alpha.m.cor$p.value, eq=FALSE)$).
%
%<<I_glm, include=FALSE>>=
%    # I_error is influenced by alpha and I but not by m
%    stopifnot(I.glm[Parameter == "alpha", min(p) < 0.05])
%    stopifnot(I.glm[Parameter == "m", min(p) > 0.05])
%    stopifnot(I.glm[Parameter == "I", min(p) < 0.05])
%
%    # correlation between I_error and true_alpha
%    I.alpha.cor <- d[,cor.test(I_error, true_alpha, method="spearman")]
%    stopifnot(I.alpha.cor$p.value < 0.05)
%
%    # correlation between I_error and true_I
%    I.I.cor <- d[,cor.test(I_error, true_I, method="spearman")]
%    stopifnot(I.alpha.cor$p.value < 0.05)
%@
%
%The mean error in the estimated value of \gls{I} was
%    d[,round(median(I_error))] 
%    [d[,round(quantile(I_error, 0.25))]-d[,round(quantile(I_error, 0.75))]],
%an over-estimate of roughly a factor of 1.5
%(\cref{fig:abcptm2,fig:abcptm1,fig:abcptm3,fig:abcptm4}). \Gls{GLM} analysis
%indicated a relationship between the error in estimated \gls{I}, and the true
%values of \gls{alpha} and \gls{I}
%    ($p$ values $pp(I.glm[Parameter == "alpha", min(p)], eq=FALSE)$
%     $pp(alpha.glm[Parameter == "I", min(p)], eq=FALSE)$,
%     \cref{tab:glmalpha}),
%but not the true value of \gls{m}. There was a significant correlation between 
%the true value of \gls{alpha} and the error in estimated \gls{I}
%    (Spearman's $\rho$ = I.alpha.cor$estimate,
%     $p pp(I.alpha.cor$p.value, eq=TRUE)$),
%and between the true value of \gls{I} and the error in estimated \gls{I}
%    (Spearman's $\rho$ = I.I.cor$estimate,
%     $p pp(I.I.cor$p.value, eq=TRUE)$).
%
%<<m_glm, include=FALSE>>=
%    # effects of parameters on m_error
%    stopifnot(m.glm[Parameter == "m", min(p)] < 0.05)
%    stopifnot(m.glm[Parameter == "alpha", min(p)] > 0.05)
%    stopifnot(m.glm[Parameter == "I", min(p)] > 0.05)
%
%    # m was only estimated correctly in just over 20\% of simulations with m > 2
%    m.tbl <- prop.table(table(d[true_m > 1,m_error]))
%    stopifnot(m.tbl[1] > 0.2 & m.tbl[1] < 0.3)
%
%    # with m = 1, over 95% of simulations are right
%    m1.est <- d[true_m == 1, m_error]
%    stopifnot(sum(m1.est == 0) / length(m1.est) > 0.95)
%    stopifnot(sum(m1.est == 0) / length(m1.est) < 1)
%@
%
%\gls{GLM} analysis showed an effect of the true value of \gls{m} on the error
%in the estimated \gls{m}
%    ($p pp(m.glm[Parameter == "m", min(p)])$,
%     \cref{tab:glmm}).
%When the true value of \gls{m} was 1, the correct value was recovered by
%\software{netabc} in virtually every case
%    (sum(m1.est == 0) out of length(m1.est) simulations).
%However, when the true value of \gls{m} was 2 or higher, the correct value was
%recovered in only 
%    as.integer(m.tbl[1] * 100)\%
%of simulations, little better than random guessing. The \gls{GLM} analysis did
%not indicate any effects of the true parameter values on the error in estimated
%\gls{m} (\cref{tab:glmm}).
%
%<<N_glm, include=FALSE>>=
%    # alpha_error is influenced by alpha and m but not I
%    stopifnot(N.glm[Parameter == "alpha", min(p) < 0.05])
%    stopifnot(N.glm[Parameter == "I", min(p) < 0.05])
%    stopifnot(N.glm[Parameter == "m", min(p) > 0.05])
%
%    N.alpha.test <- d[,wilcox.test(.SD[true_alpha == 1.5, N_error], 
%                                   .SD[true_alpha < 1.5, N_error])]
%    stopifnot(N.alpha.test$p.value < 0.05)
%    N.I.test <- d[,wilcox.test(.SD[true_I == 3000, N_error], 
%                               .SD[true_I < 3000, N_error])]
%    stopifnot(N.I.test$p.value < 0.05)
%@
%
%Finally, the total number of nodes \gls{N} was consistently over-estimated by
%about a factor of two
%    (error d[,format(median(N_error), scientific=FALSE)] 
%    [d[,format(quantile(N_error, 0.25), scientific=FALSE)] - 
%     d[,format(quantile(N_error, 0.75), scientific=FALSE)]]).
%The fitted \gls{GLM} indicated that the true values of both \gls{alpha} and
%\gls{I} had an effect on the error in the estimated \gls{N}
%    ($p$-values $pp(N.glm[Parameter == "alpha", min(p)], eq=FALSE)$ and 
%     $pp(N.glm[Parameter == "I", min(p)], eq=FALSE)$,
%     \cref{tab:glmN}),
%but that the true value of \gls{m} did not. The error in the estimated \gls{N}
%when \gls{alpha} was equal to 1.5 was slightly lower than for other values of
%\gls{alpha}
%    (median [IQR] error rates 
%     d[true_alpha == 1.5, format(median(N_error), scientific=FALSE)] 
%    [d[true_alpha == 1.5, format(quantile(N_error, 0.25), scientific=FALSE)] - 
%     d[true_alpha == 1.5, format(quantile(N_error, 0.75), scientific=FALSE)]]
%     for \gls{alpha} = 1.5 vs. 
%     d[true_alpha < 1.5, format(median(N_error), scientific=FALSE)] 
%    [d[true_alpha < 1.5, format(quantile(N_error, 0.25), scientific=FALSE)] - 
%     d[true_alpha < 1.5, format(quantile(N_error, 0.75), scientific=FALSE)]]
%     for \gls{alpha} < 1.5),
%and this difference was statistically significant
%    (Wilcoxon rank-sum test, $p pp(N.alpha.test$p.value, eq=TRUE)$).
%Similarly, the error in the estimated \gls{N} when \gls{I} was equal to 3000 was
%lower than for other values of \gls{I}
%    (median [IQR] error rates 
%     d[true_I == 3000, format(median(N_error), scientific=FALSE)] 
%    [d[true_I == 3000, format(quantile(N_error, 0.25), scientific=FALSE)] - 
%     d[true_I == 3000, format(quantile(N_error, 0.75), scientific=FALSE)]]
%     for \gls{I} = 3000 vs. 
%     d[true_alpha < 3000, format(median(N_error), scientific=FALSE)] 
%    [d[true_alpha < 3000, format(quantile(N_error, 0.25), scientific=FALSE)] - 
%     d[true_alpha < 3000, format(quantile(N_error, 0.75), scientific=FALSE)]]
%     for \gls{I} < 3000),
%and this difference was statistically significant
%    (Wilcoxon rank-sum test, $p pp(N.I.test$p.value, eq=TRUE)$).
%
%\begin{figure}
%  \includegraphics{abc-point-estimate-m2}
%  \caption[\Acrlong{MAP} point estimates for \gls{BA} model parameters obtained
%    by running \software{netabc} on simulated data, for simulations with $m = 2$.] 
%  {
%    \Acrlong{MAP} point estimates for \gls{BA} model parameters obtained by         
%    running \software{netabc} on simulated data. Values shown are for               
%    simulations with \gls{m} = 2. Dashed lines indicate true values. (A)            
%    Estimates of \gls{alpha} and \gls{I} which were varied in these simulations  
%    against known values. (B) Estimates of \gls{m} and \gls{N} which were held   
%    fixed in these simulations at the values \gls{m} = 2 and \gls{N} = 5000. 
%  }
%  \label{fig:abcptm2}
%\end{figure}
%
%<<posterior_sims, echo=FALSE>>=
%    N <- 5000
%    replicate <- 0
%    lab <- NULL
%    for (m in c(2, 3, 4)) {
%    for (I in c(1000, 2000)) {
%    for (alpha in c(0, 0.5, 1, 1.5)) {
%        lab <- c(lab, sprintf("fig:%.1f-%d-%d-%d-%d", alpha, I, m, N, replicate))
%    }
%    }
%    }
%    lab <- paste(lab, collapse=",")
%@
%
%\Cref{fig:abcex} shows the \gls{ABC} approximation to the posterior
%distribution on the \gls{BA} parameters for one simulation. Equivalent plots
%for one replicate simulation with each combination of parameters can be found
%in \cref{lab}. \Gls{HPD} intervals around \gls{alpha} and \gls{I} were
%narrow relative to the region of nonzero prior density, whereas the intervals
%for $m$ and \gls{N} were widely dispersed. \Cref{tab:abchpd} shows point
%estimates and 95\% \gls{HPD} intervals averaged over all simulations.
%
%\begin{figure}
%  \includegraphics{{abc-posterior/1.0_1000_2_5000_0}.pdf}
%  \caption[Approximate marginal posterior distributions of BA model parameters
%      obtained by applying \textit{netabc} to a simulated transmission tree
%      with values $\alpha$ = 1.0, $I$ = 1000, $m$ = 2, and $N$ = 5000.]
%    {
%        Approximate marginal posterior distributions of BA model parameters
%        obtained by applying \textit{netabc} to a simulated transmission tree
%        with BA parameter values $\alpha$ = 1.0, $I$ = 1000, $m$ = 2, and $N$ =
%        5000. Vertical dashed lines indicate true values. Shaded areas are 95\%
%        highest posterior density intervals. $x$-axes indicate regions of
%        nonzero prior density.
%    }
%  \label{fig:abcex}
%\end{figure}
%
%
%\begin{table}
%    \centering
%    \input{\tablepath/abc-hpd}
%    \caption[
%        Maximum \textit{a priori} estimates and 95\% highest posterior density
%        (HPD) interval boundaries for \gls{BA} model parameters estimated with
%        \software{netabc}, averaged over simulated transmission trees.
%    ]
%    {
%        Maximum \textit{a priori} estimates and 95\% highest posterior density
%        (HPD) interval boundaries for \gls{BA} model parameters estimated with
%        \software{netabc}, averaged over simulated transmission trees.
%    }
%    \label{tab:abchpd}
%\end{table}
%
%<<mixed_peerdriven, include=FALSE>>=
%  library(coda)
%  d <- fread("bzcat ../../simulations/abc-pa-mixed-alpha/abc/*")
%  d <- d[iter == max(iter)]
%  d <- d[sample(1:nrow(d), prob=weight, replace=TRUE)]
%  params <- c("alpha", "m", "N", "I")
%  d <- melt(d, measure.vars=params, variable.name="parameter")
%  f <- function (x) {
%    dens <- density(x)
%    hpd <- HPDinterval(mcmc(x))
%    list(map=dens$x[which.max(dens$y)], hpd.min=hpd[,"lower"], 
%         hpd.max=hpd[,"upper"])
%  }
%  d <- d[,f(value), by=parameter]
%  setkey(d, parameter)
%
%  m <- collect.metadata("../../simulations/abc-pa-peerdriven/abc/*")
%  p <- fread(paste("bzcat", rownames(m)[m$sample_peer == 1]))
%  p <- p[iter == max(iter)]
%  p <- p[sample(1:nrow(p), prob=weight, replace=TRUE)]
%  params <- c("alpha", "m", "N", "I")
%  p <- melt(p, measure.vars=params, variable.name="parameter")
%  f <- function (x) {
%    dens <- density(x)
%    hpd <- HPDinterval(mcmc(x))
%    list(map=dens$x[which.max(dens$y)], hpd.min=hpd[,"lower"], 
%         hpd.max=hpd[,"upper"])
%  }
%  p <- p[,f(value), by=parameter]
%  setkey(p, parameter)
%  options(scipen=5)
%@
%
%To test the effect of model misspecification, we simulated one network where
%the nodes exhibited heterogeneous preferential attachment power (half 0.5, the
%other half 1.5), with $m$ = 2, $N$ = 5000, and $I$ = 1000. The MAP [95\%
%HPD] estimates for each parameter were: 
%$\alpha$, 
%  round(d["alpha", map], 2) 
%  [round(d["alpha", hpd.min], 2) -
%   round(d["alpha", hpd.max], 2)];
%$I$,
%  d["I", round(map)] 
%  [d["I", round(hpd.min)] -
%   d["I", round(hpd.max)]];
%$m$,
%  d["m", floor(map)] 
%  [d["m", floor(hpd.min)] -
%   d["m", floor(hpd.max)]];
%$N$,
%  d["N", round(map)] 
%  [d["N", round(hpd.min)] -
%   d["N", round(hpd.max)]].
%The approximate posterior distributions for this simulation are shown in
%\cref{fig:mixed}. To test the effect of sampling bias, we sampled one
%transmission tree in a peer-driven fashion, where the probability to sample a
%node was twice as high if one of its peers had already been sampled. The
%parameters for this experiment were $N$ = 5000, $m$ = 2, $\alpha$ = 0.5, and
%$I$ = 2000. The estimated values were
%$\alpha$, 
%  p["alpha", round(map, 2)] 
%  [p["alpha", round(hpd.min, 2)] -
%   p["alpha", round(hpd.max, 2)]];
%$I$,
%  p["I", floor(map)] 
%  [p["I", floor(hpd.min)] -
%   p["I", floor(hpd.max)]];
%$m$,
%  p["m", floor(map)] 
%  [p["m", floor(hpd.min)] -
%   p["m", floor(hpd.max)]];
%$N$,
%  p["N", floor(map)] 
%  [p["N", floor(hpd.min)] -
%   p["N", floor(hpd.max)]].
%The approximate posterior distributions are shown in \cref{fig:peerdriven}. Both
%of these results were in line with estimates obtained on other simulated
%datasets (\cref{tab:abchpd}), although the estimate of peer-driven sampling for
%$\alpha$ was somewhat lower than typical.

@@

To make our analyses comparable to the existing network literature, we
estimated values of the power law exponent $\gamma$ for each of the datasets 
investigated. The results are summarized in \cref{tab:gamma}. All of the
estimated exponents were in the range $2 \leq \gamma \leq 4$ commonly reported
in the literature~\autocite{liljeros2001web, schneeberger2004scale,
colgate1989risk, brown2011transmission}.
